---
layout: post
title: "How to View Source Code in R"
date: 2014-10-09
comments: true
categories: r
keywords: "R, source code of a function"
published: true
share: true
ads: true
---

The best way to learn how to use a function written by other people is to read its source code. If you use Rstudio, you just need to type the function name at the command line. However, a lot of times, you will get something like this. 
```{r}
mean
```

You may wonder, "where is the function body?" You don't see the function body because `mean()` is a **generic function**. Generic functions are used to implement the S3 object oriented system in R. Each generic function has one or many methods. For example, `mean()` has the following methods:
```{r}
methods(mean)
```

We cannot view the source code of a generic function, instead, we can view the source code of each generic.method function. For example, run the following commands to view the source of `mean.default()` and `mean.Date()` respectively.
```{r}
getAnywhere(mean.default)
getAnywhere(mean.Date)
```

