---
layout: post
title: "How to Write Infix Functions in R"
date: 2014-09-05 15:18:35 -0400
comments: true
categories: r
keywords: "R, infix functions"
published: true
share: true

---
Infix functions are functions where the function name comes in between its arguments when used. For example, operators like + and - are actually infix functions. Conversely, you can think of infix functions as operators. Here's an example of an infix function that concatenate two vectors.
```{r}
`%&%` <- function(x, y) { 
        # @x, @y: vectors or sigle values
        # returns: this function concatenates @x and @y without space in between
        paste(x, y, collape="", sep="")
}
```

To use it, we simply treat it as any other operators.
```{r}
"data" %&% "analysis"
"cabaceo" %&% 1:5
```

Here's another example of an infix function that allows you to return a default value when encountering NULL. This is very helpful when used inside of another function to avoid returning a NULL value.
```{r}
`%||%` = function(a, b) if (!is.null(a)) a else b

100 %||% NA
NULL %||% NA
```

To learn more, read [the chapter on functions](http://adv-r.had.co.nz/Functions.html) from Hadley's book, Advanced R.