---
layout: post
title: "A Super Easy Way to Make Ggplot2 Bar Charts with Labels, ezplot - Part 4"
date: 2015-05-22 
comments: true
categories: r
keywords: "R, ggplot, ggplot2, ezplot, bar plots, bar charts, bar charts with labels, label the bars, bars with labels"
published: true
share: true
ads: true
---

### Updated October 4, 2018

Previously, I showed how to use the `mk_barplot_resp()` function to make [bar charts](https://masterr.org/r/an-easy-way-to-make-ggplot2-plots-ezplot-part1/). 
In this post, I'll show you how to use the `mk_barplot_freq()` function.

Make sure you first install ezplot by running the command `devtools::install_github("gmlang/ezplot")`.
```{r, warning=FALSE, message=FALSE}
library(ezplot)
library(dplyr)
```

We'll use the diamonds data in the ggplot2 package. 
```{r, warning=FALSE, message=FALSE}
plt = mk_barplot_freq(ggplot2::diamonds)
```

Let's make a bar chart to show the frequency distribution of the clarities.
Notice the bars are automatically labeled with both counts and percents.
```{r, warning=FALSE, message=FALSE}
plt("clarity")
```

We can also choose to show relative frequencies (%) on the y-axis. Just set
`show_pct = TRUE`
```{r, warning=FALSE, message=FALSE}
plt("clarity", show_pct = TRUE)
```

If [ezplot](https://leanpub.com/ezplot) has improved your productivity, please 
share it with your friends.
