---
layout: post
title: "An Easy Way to Make Ggplot2 Boxplots, ezplot - Part 2"
date: 2015-04-26 12:14:49 -0400
comments: true
categories: R
keywords: "R, ggplot, ggplot2, ezplot, boxplot"
published: true
share: true

---

Previously, I introduced the ezplot package and [demoed how to use it to make nice looking ggplot2 bar charts easily and quickly](http://masterr.org/r/an-easy-way-to-make-ggplot2-plots-ezplot-part1/).   Today, I'm going to show you how to make nice looking boxplots. Once again, the code is super simple.

### Install and Load ezplot
```{r, warning=FALSE, message=FALSE}
# First install devtools if you don't already have it installed, with:
# install.packages("devtools")
devtools::install_github("gmlang/ezplot")
library(ezplot)
```

### The ezplot package comes with a films dataset obtain from IMBD.com, and I'll use it to demo boxplots. First, let's draw boxplots of budget vs. years. 
```{r, warning=FALSE, message=FALSE}
# call the films dataset make a function that will draw boxplots 
# using the variables in the films data set
plt = mk_boxplot(films)

# plot distributions of budget over the years
title1 = "Annual Distribution of Budget from 1913 to 2014"
p = plt("year_cat", "budget", ylab="budget", main=title1)
print(p)
```

We see the y-axis tick labels are expressed in scientific notations. This makes it difficult to read. We can use the function `r scale_axis()` to use the comma scale on the y-axis, which will display the numbers in 000,000 format.
```{r}
scale_axis(p, "y", use_comma=T)
```

We also observe that all the boxes are squashed down, indicating budget is heavily right-skewed. We can use either the log scale or the log10 scale on the y-axis. Once again, this can be easily done using the function `r scale_axis()`.
```{r}
scale_axis(p, "y", use_log=T)
scale_axis(p, "y", use_log10=T)
```

I created [ezplot](https://github.com/gmlang/ezplot). Drop a comment below if you have questions.