---
layout: post
title: "How to easily make beautiful heatmaps with ezplot - Part 8"
date: 2015-07-30 
comments: true
categories: r
keywords: "R, ezplot, ggplot, ggplot2, heatmap, heat map, NBA data analysis, NBA statistics, Dwyane Wade, LeBron James, Dirk Nowitzki, Kevin Durant, Koby"
published: true
share: true
ads: true
---

### Updated October 4, 2018

In this post, we'll look at how to make effective heatmaps using ezplot. We'll 
use a dataset of NBA players' statistics from flowingdata.com. Make sure you 
first install ezplot by running the command `devtools::install_github("gmlang/ezplot")`.
```{r, warning=FALSE, message=FALSE}
library(ezplot)
library(dplyr)
library(tidyr)
```

Let's get the data. Notice we pass the url directly to `read.csv()`.
```{r, warning=FALSE, message=FALSE}
nba = read.csv("http://datasets.flowingdata.com/ppg2008.csv")
# examine the variables
str(nba)
# look at the first 5 rows and 8 columns
nba[1:5, 1:8]
```

The variable `Name` has the names of the NBA players. By default, it's treated 
by R as a Factor with levels ordered alphabetically. Reorder its levels by 
points scored.
```{r, warning=FALSE, message=FALSE}
nba$Name = with(nba, reorder(Name, PTS))
```

The other variables are various performance statistics. Before we can make a 
heat map, we need to put the data in the long format. In other words, we need to 
gather the names of the statistics in one column, and their values in another column.
```{r, warning=FALSE, message=FALSE}
nba_m = nba %>% gather(stats, val, -Name)
head(nba_m)
```

We also want to scale the values of each performance statistics so that they are 
between 0 and 1. 
```{r, warning=FALSE, message=FALSE}
dat = nba_m %>% group_by(stats) %>% mutate(val_scaled = scales::rescale(val))
head(dat)
```

With all these data prep work done, we're ready to make a heatmap. This is super 
easy with ezplot.
```{r, warning=FALSE, message=FALSE}
plt = mk_heatmap(dat)
p = plt("stats", "Name", "val_scaled")
rotate_axis_text(p, 90, vjust_x = 0.5)
```

Not it's your turn. Make a heat map using the unscaled values, and compare it 
with the scaled version. You will see they are very different. The scaled 
version is the mathematically correct one.

This is the last post in the ezplot how-to series. If you've enjoyed it, 
tell your friend about it. If you want to learn more about how to use
ezplot, you can get my book [here](https://leanpub.com/ezplot). 

