---
layout: post
title: "How to Scrape Japanese Text Using the rvest Package"
date: 2015-01-07 10:04:49 -0400
comments: true
categories: R
keywords: "R, web scraping, Japanese characters, rvest, html"
published: true
share: true

---

The [rvest](https://github.com/hadley/rvest) package is yet another 
powerful tool developed by Hadley. It makes web scraping using R easy, however, sometimes, it fails on non-English languages. For example, say I want to scrape [this page](http://www3.boj.or.jp/market/jp/stat/of141205.htm) from the Bank of Japan. I can try to use the `html` function directly while specifying `encoding="UTF-8"`, but this will not display the Japanese character correctly.

```{r, warning=FALSE, message=FALSE}
library(rvest)
url = "http://www3.boj.or.jp/market/jp/stat/of141205.htm"
page = html(url, encoding="UTF-8")

# extract title
page %>% html_nodes("title") %>% html_text()

# extract 2nd table
page %>% html_nodes('table') %>% .[[2]] %>% html_table()
```

To solve this issue, we can use `readLines()` and then `html()`.
```{r, warning=FALSE, message=FALSE}
page = html(readLines(url, encoding = "UTF-8"))

# extract title
page %>% html_nodes('title') %>% html_text()

# extract 2nd table
page %>% html_nodes('table') %>% .[[2]] %>% html_table()
```
