---
layout: post
title: "Examples of Predicate Functionals"
date: 2016-03-19 
comments: true
categories: r
keywords: "R, predicate functional, Filter, where, Find, Position"
published: true
share: true
ads: true
---

A predicate is a function that returns a single TRUE or FALSE, for example, `is.factor()`, `all()`, or `is.NULL()`. A predicate functional is a function that applies a predicate to each element of a list or data frame. For example, we can define a predicate functional `where()` that checks the type of each column in a data frame.
```{r, warning=FALSE, message=FALSE}
where = function(f, x) {
        # Applies a predicate function to every element of a list or data frame.
        # f: a predicate function
        # x: a list or data frame
        #
        # Returns a logical vector.
        vapply(x, f, logical(1))
}
```

Here's how you'd use it.
```{r, warning=FALSE, message=FALSE}
# make a example data frame
df = data.frame(x = 1:3, y = c("a", "b", "c"), z = c("Alex", "Bob", "Chris"))

# check if each column of the data frame is a factor
where(is.factor, df)

# check if each column of the data frame is an integer
where(is.integer, df)

# check if each column of the data frame is a character
where(is.character, df)
```

There're 3 common predicate functionals already defined in base R, namely, `Filter(), Find(), Position()`. Here's how you can use them.
```{r, warning=FALSE, message=FALSE}
# filter the factors in a data frame
subdf = Filter(is.factor, df)
subdf
str(subdf) # note: Filter() returns a data frame

# find the 1st factor starting from the Left in a data frame
x_factor = Find(is.factor, df)
x_factor
str(x_factor) # note: Find() returns the first factor it finds from the left

# gives the position of the 1st factor starting from the Right in a data frame
Position(is.factor, df, right=T)
```

This article is inspired by Hadley's book "Advanced R", you can <a rel="nofollow" href="http://www.amazon.com/gp/product/1466586966/ref=as_li_tl?ie=UTF8&camp=1789&creative=9325&creativeASIN=1466586966&linkCode=as2&tag=cabaceo-20&linkId=2GDWMZSF4NX32QIO">buy it from Amazon</a><img src="http://ir-na.amazon-adsystem.com/e/ir?t=cabaceo-20&l=as2&o=1&a=1466586966" width="1" height="1" border="0" alt="" style="border:none !important; margin:0px !important;" />
or read it [online](http://adv-r.had.co.nz/Functions.html) for FREE.