---
layout: post
title: "The on.exit() Function"
date: 2017-03-03
comments: true
categories: r
keywords: "R, on.exit"
published: true
share: true
ads: true
---

There's an interesting little function in R called `on.exit()`. It can be used 
in your own function to perform some side effect. For example, in addition to 
returning a value, the following function uses `on.exit()` to also print two messages.
```{r, warning=FALSE, message=FALSE}
myfun = function(x){
        on.exit(print("first"))
        on.exit(print("second"), add = TRUE)
        return(x)
}
myfun(2)
```

Note what happens if we remove `add=TRUE` from the second `on.exit()` usage.
```{r, warning=FALSE, message=FALSE}
fun = function(x){
        on.exit(print("first"))
        on.exit(print("second"))
        return(x)
}
fun(2)
```
We see "first" isn't printed anymore.

To learn more about `on.exit()`, refer to Hadley's book "Advanced R." You can <a rel="nofollow" href="http://www.amazon.com/gp/product/1466586966/ref=as_li_tl?ie=UTF8&camp=1789&creative=9325&creativeASIN=1466586966&linkCode=as2&tag=cabaceo-20&linkId=2GDWMZSF4NX32QIO">buy it from Amazon</a><img src="http://ir-na.amazon-adsystem.com/e/ir?t=cabaceo-20&l=as2&o=1&a=1466586966" width="1" height="1" border="0" alt="" style="border:none !important; margin:0px !important;" />
or read it [online](http://adv-r.had.co.nz/Functions.html) for FREE.