---
layout: post
title: "RFP - Part2: R Lists"
date: 2019-02-12
comments: true
categories: r
keywords: "R, lists, R data types, R data structures, recursion in R, RFP series, R functional programming, R is a functional programming language, FP, functional programming in R, head, tail, is empty in R, check a R list is empty, R tuples, R pairs, tuples in R, pairs in R, R lists are vectors"
published: true
share: true
ads: true
---

Previously, we discussed R vectors. We now turn to R lists. Like a vector, a R list can also grow or shrink. But unlike a vector, a R list can contain any R objects. For example, vectors, lists, functions, or environments can all be the elements of a list, and it's perfectly okay to mix them in the same list. 

## The empty list

The empty list, with syntax `list()`, has 0 elements. Indeed, if we call `length(list())`, we get 0.

## Non-empty lists

We can make a list with `list(e1, ..., en)` where each expression[^1] `e` is evaluated to a R object. It's more common to make a list with `c(e1, e2)`, called called "`e1` combined with `e2`," where `e1` evaluates to a "list" and `e2` evaluates to another "list." The result is a new list that starts with the elements in `e1` followed by the elements in `e2`. In the simpler case when `e1` evaluates to a single value like 5, 1.2, "hello", TRUE or NA, `c(e1, e2)` can also be called "`e1` consed onto `e2`." The result is then a new list that starts with the value of `e1` followed by the elements in `e2`. Here're some examples of lists:

```{r}
list(1L, 2L, 3L) # each element is an integer
list(c(1, 2), c("gm", "lang")) # each element is a length-2 vector of diff. type
list(list(1, 2), TRUE, function(x) x+1, environment()) # 1st element is a list, 3rd element is a function, 4th element is an env object

# cons 9 onto the list(3, 6)
c(9, list(3, 6)) 

# list(1, list(2, 3)) is different from list(1, 2, 3)
!identical(list(1, list(2, 3)), list(1, 2, 3))

# list(1, list(2, 3)) is different from list(list(1, 2), 3))
!identical(list(1, list(2, 3)), list(list(1, 2), 3))
```

## How to use lists

Once again, powered by recursion, we only need three simple operations when working with lists:

1. Check if a list is empty. 
2. Get the first element of a list, raising an exception if the list is empty.
3. Get the tail of a list without its first element, raising an exception if the list is empty.

Like in the case of vectors, R doesn't provide these operations perfectly out of the box. So we need to write them ourselves:
```{r}
# Returns TRUE if a list is empty, FALSE otherwise.
is_empty = function(xs) { # xs: a list 
        length(xs) == 0
}

# Get the first element of a list, raising an exception if the list is empty.
hd = function(xs) { # xs: a list 
        if (is_empty(xs)) stop("List is empty.")
        else xs[[1]] 
}

# Get the tail of a list without its first element, raising an exception if the list is empty.
tl = function(xs) { # xs: a list 
        if (is_empty(xs)) stop("List is empty.")
        else xs[-1] 
}
```

Having defined `is_empty()`, `hd()` and `tl()`, we can use them inside of recursive functions to perform complex operations on lists. For example, given a list of integer pairs, we can write a function to sum them up. 
```{r}
sum_pair_list = function(xs) {
        # xs: a list of interger pairs represented by length-2 vectors c(i1, i2)
        if (is_empty(xs)) 0
        else hd(xs)[1] + hd(xs)[2] + sum_pair_list(tl(xs))
}
sum_pair_list(list())
sum_pair_list(list(c(1, 2), c(3, 4)))
```

Or we can write functions to put the first or second element of each pair into separate lists.
```{r}
firsts = function(xs) {
        # xs: a list of interger pairs represented by length-2 vectors c(i1, i2)
        if (is_empty(xs)) list()
        else c(hd(xs)[1], firsts(tl(xs)))
}

seconds = function(xs) {
        # xs: a list of interger pairs represented by length-2 vectors c(i1, i2)
        if (is_empty(xs)) list()
        else c(hd(xs)[2], seconds(tl(xs)))
}

xs = list(c(1, 2), c(3, 4))
firsts(xs)
seconds(xs)
```

It should make us cringe that `firsts()` and `seconds()` look almost identical yet we wrote them as two different functions. We'll learn how to fix it later. Notice that `is_empty()`, `hd()` and `tl()` defined above are very similarly to the ones defined in [part1 of this series when we discussed R vectors](https://masterr.org/r/RFP-part1-vectors/). This is because lists are just a special kind of vectors in R. To see this, realize that another way of declaring an empty list is `vector("list")`.
```{r}
identical(vector("list"), list())
```


[^1]: The word 'expression' here is being used in the sense of expressions in any programming language or mathematical expressions. Do NOT confuse it with the R expression object. 
